Supporting information: Copulas and their potential for ecology

### 2019-09-10 Paper.Rmd
still working on Background section

All branches of ecology 
study relationships among environmental and biological 
variables. However, commonly used correlation and regression 
approaches to studying 
such relationships are limited, and 
provide only a partial description of a relationship between variables. 
For instance, datasets showing markedly different 
relationships may have the same correlation coefficient
(Fig. \ref{fig_cop_pedag1}A-B). The variables of Fig. \ref{fig_cop_pedag1}A 
(respectively, Fig. \ref{fig_cop_pedag1}B) are principally 
related in the left (respectively, right) portions of their distributions, 
an asymmetric pattern of dependence that can
have ecological significance, as discussed below, but is not captured by 
correlation. Although correlation is only one way to 
study relationships between variables, other common metrics also 
provide only partial information.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
\caption{(A-D) Bivariate datasets showing diverse relationships between 
variables; some of these datasets nevertheless have the same Pearson (P)
or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
respectively. (E,F) Reflections
of C, D, respectively, about a vertical axis. Subscripts h and v 
in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
\end{center}
\end{figure}
<!--***DAN: When I cut panels out of this figure (saving them for use in a 
later figure), I also cut the caption, and here is the cut stuff for later use:
(C, D) Contours 
of probability density functions (pdf) of Clayton (C) and survival-Clayton 
(D) copulas with parameter 5 (see section on methods for Q1), together
with samples from the corresponding distributions. (E,F) Reflections
of C, D, respectively, about a vertical axis. (G) Contours of the
pdf of a normal copula (methods for Q1) with 
parameter chosen to
have the same Spearman correlation as A-D, together with a sample.
(H) The portion of the Spearman correlation within a band 
can be computed
for any band (methods for Q1) to describe how dependence
varies from one part of the two distributions to another. Diagonal lines
show two bands, the data in the left/lower band showing stronger
dependence than the right/upper band.
Subscripts h and v 
stand for "horizontal" and "vertical".
-->

@@

A main benefit of a copula approach is 
that it can detect tail dependence and asymmetries of tail dependence. 
Tail dependence (defined precisely in the section on Background, below) 
is association between extreme values of variables. 
If smaller values of two positively associated variables are more 
strongly associated than are larger values, the variables are said to
have stronger lower- or left-tail dependence than right- or upper-tail dependence
(Fig. \ref{fig_cop_pedag1}C);
and *vice versa* if larger values are more strongly associated than smaller
ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
Spearman or Kendall correlation can have a range of tail dependence 
(Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
methods for Q1 below). Although copulas can be used to model the entire
complex dependence structure of data, we have found tail dependence to be
a useful aspect of the full structure of dependence, so tail dependence 
is one of the main foci of this paper.

@@

We approached our overall goal of exploring the potential of copula approaches 
for ecology by addressing the following specific questions. (Q1) Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do they show tail 
dependence distinct from that of a normal distribution, and in particular
do they show asymmetric tail dependence? Normal copula structure is assumed
by standard approaches that use multivariate 
normal distributions or distributions obtained by transforming
the marginals of a normal distribution. So Q1 asks whether ecological
data contain dependency information distinct from what can be modelled with
standard approaches. (Q2) If the answer to
Q1 is "yes," what are some possible causes/mechanisms of non-normal
copula structure and tail dependence? And (Q3) what are the 
consequences of non-normal copula
structure and tail dependence for ecological understanding and
applications? After a Background section (the second section), we address Q1 
(third through fifth sections of the paper) 
by analyzing several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data, selected to span multiple 
sub-fields and organizational levels of ecology. 
We address Q2 (sixth and seventh sections of the paper), using 
simple models. We address Q3 (eighth and ninth sections of the paper) using 
both data and models. The final section is the Discussion.
Multiple analyses are brought to bear on each question, and these
are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
serve as a *post hoc* guide to the paper. 
Copula approaches have been used to great effect in neuroscience [@onken2009], 
bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
but have been used only rarely so far in ecology (but see @Valpine2014, 
@Anderson2018 and @Popovic2018). 
We argue that benefits wider usage of complete 
copula descriptions of dependence in ecology will be substantial.

@@

We give a brief introduction to copulas, focusing, for simplicity, on bivariate copulas
and on concepts needed
for the rest of the paper. See @nelsen2006_copula, @joe2014_dependence, and
@MaiScherer2017 for general background, and see also @Anderson2018, part of 
which is a short introduction to 
copulas for ecologists. A bivariate *copula* can be defined as a bivariate distribution
function with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
This will be the
distribution function of a bivariate random variable with uniform marginals, and 
the terminology *copula* is sometimes also applied to the random variable. 
A foundational theorem of
@sklar1959_theorem (see also @MaiScherer2017, p. 16) 
says that if $F$ is the bivariate distribution function
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
exists a copula $C$
such that for all $(x,y)$ in the Euclidean plane, $F(x,y)=C(F_X(x),F_Y(y))$.
The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
Thus $C$ couples the bivariate distribution, $F$, with the two marginal distributions. 
Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
$G_A$ and $G_B$ are univariate distribution functions, then 
$D(G_A(x),G_B(y))$ is a distribution function.
The applications of this study fall within the case in which univariate marginal 
distributions have continuous, strictly 
monotonic distribution functions, and this case is simpler, so we 
henceforth make such an assumption. 
Then, "the" copula (now guaranteed unique) 
associated with $(X,Y)$ is the distribution
function of the random variable $(U,V)=(F_X(X),F_Y(Y))$ 
(Appendix \ref{ContStrictMonoton}). And 
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
vector with distribution function $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
(Appendix \ref{ContStrictMonoton}).

Thus Sklar's theorem implies that
any random vector can be constructed from a copula and marginal 
distributions. And conversely,
any copula and any univariate distribution functions give rise to a random vector. 
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables. Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure. The theorem
also make it possible to study and model dependence independently 
from marginal distributions. For instance, in a pedagogical 
example, [@Anderson2018] modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then separately modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].
<!--***DAN: dont forget to put these few words in!-->

<!--***DAN: I still feel like we have to make this more intuitive for people
at this point. Maybe a figure with two panels showing two copulas on
the left, and two panels shwoing two sets of marginals at the top,
and then four panels showing what you get when each of these combine?
Good idea. Steps: 
DONE split the current Fig 1 so the new Fig 1 has the first 6 panels,
but C-D should only have the points. 
Then make a Fig 2 as above, and edit the text in the above para to use that figure. 
The rest of the panels of the current Fig 1 will make a new Fig 3 later.
-->

<!--***DANr: Add a new section called Background on copulas.
It can have the following kinds of stuff (probably not exhaustive, and probably
not in order, and some of this material is currently below):
1) DONE - The definition, Sklar's theorem
2) What good are copulas - intuitive idea, basically, but also have to be linked to 
the formalities of the definition

What some of the different copula families look like
The explanation of how ranks relate to copulas and estimating the copula parameters
How to carry out fitting (basically, MLE fitting on ranks - Genest et al paper)
What tail dependence actually is
-->

@@

@@

>>>>

All branches of ecology 
study relationships among environmental and biological 
variables. However, commonly used correlation and regression 
approaches to studying 
such relationships are limited, and 
provide only a partial description of a relationship between variables. 
For instance, datasets showing markedly different 
relationships may have the same correlation coefficient
(Fig. \ref{fig_cop_pedag1}A-B). The variables of Fig. \ref{fig_cop_pedag1}A 
(respectively, Fig. \ref{fig_cop_pedag1}B) are principally 
related in the left (respectively, right) portions of their distributions, 
an asymmetric pattern of dependence that can
have ecological significance, as discussed below, but that is not captured by 
correlation. Although correlation is only one way to 
study relationships between variables, other common metrics also 
provide only partial information.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
\caption{(A-D) Bivariate datasets showing diverse relationships between 
variables; some of these datasets nevertheless have the same Pearson (P)
or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
respectively. (E,F) Reflections
of C, D, respectively, about a vertical axis. Subscripts h and v 
in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
\end{center}
\end{figure}
<!--***DAN: When I cut panels out of this figure (saving them for use in a 
later figure), I also cut the caption, and here is the cut stuff for later use:
(C, D) Contours 
of probability density functions (pdf) of Clayton (C) and survival-Clayton 
(D) copulas with parameter 5 (see sections on methods for Q1), together
with samples from the corresponding distributions. (E,F) Reflections
of C, D, respectively, about a vertical axis. (G) Contours of the
pdf of a normal copula (methods for Q1) with 
parameter chosen to
have the same Spearman correlation as A-D, together with a sample.
(H) The portion of the Spearman correlation within a band 
can be computed
for any band (methods for Q1) to describe how dependence
varies from one part of the two distributions to another. Diagonal lines
show two bands, the data in the left/lower band showing stronger
dependence than the right/upper band.
Subscripts h and v 
stand for "horizontal" and "vertical".
-->

@@

A main benefit of a copula approach is 
that it can detect tail dependence and asymmetries of tail dependence. 
Tail dependence (defined precisely in the section on Background, below) 
is association between extreme values of variables. 
If smaller values of two positively associated variables are more 
strongly associated than are larger values, the variables are said to
have stronger lower- or left-tail dependence than right- or upper-tail dependence
(Fig. \ref{fig_cop_pedag1}C);
and *vice versa* if larger values are more strongly associated than smaller
ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
Spearman or Kendall correlation can have a range of tail dependence 
(Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see sections on Background and
methods for Q1 below). Although copulas can be used to model the entire
complex dependence structure of data, we have found tail dependence to be
a useful aspect of the full structure of dependence, so tail dependence 
is one of the main foci of this paper.

@@

We approached our overall goal of exploring the potential of copula approaches 
for ecology by addressing the following specific questions. (Q1) Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do they show tail 
dependence distinct from that of a normal distribution, and in particular
do they show asymmetric tail dependence? Normal copula structure is assumed
by standard approaches that use multivariate 
normal distributions or distributions obtained by transforming
the marginals of a normal distribution. So Q1 asks whether ecological
data contain dependency information distinct from that standard or default cases. 
(Q2) If the answer to
Q1 is "yes," then what are some possible causes/mechanisms of non-normal
copula structure and tail dependence in ecology? And (Q3) what are the 
consequences of non-normal copula
structure and tail dependence for ecological understanding and
applications? After a Background section (the second section), we address Q1 
(third through fifth sections of the paper) 
by analyzing several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data, selected to span multiple 
sub-fields and organizational levels of ecology. 
We address Q2 (sixth and seventh sections of the paper), using 
simple models. We address Q3 (eighth and ninth sections of the paper) using 
both data and models. The final section is the Discussion.
Multiple analyses are brought to bear on each question, and these
are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
serve as a *post hoc* guide to the paper. 
Copula approaches have been used to great effect in neuroscience [@onken2009], 
bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
and have been used very effectively, but rarely so far in ecology [@Valpine2014; 
@Anderson2018; @Popovic2018]. 
We argue that benefits of wider usage of complete 
copula descriptions of dependence in ecology will be substantial.

@@

We give a brief introduction to copulas, focusing, for simplicity, on bivariate copulas
and on concepts needed
for the rest of the paper. See @nelsen2006_copula, @joe2014_dependence, and
@MaiScherer2017 for general background, and see also @Anderson2018, part of 
which is a short introduction to 
copulas for ecologists. A bivariate *copula* can be defined as a bivariate distribution
function with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
This will be the
distribution function of a bivariate random variable with uniform marginals, and 
the terminology *copula* is sometimes also applied to the random variable. 
A foundational theorem of
@sklar1959_theorem (see also @MaiScherer2017, p. 16) 
says that if $F$ is the bivariate distribution function
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
exists a copula $C$
such that for all $(x,y)$ in the Euclidean plane, $F(x,y)=C(F_X(x),F_Y(y))$.
The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
Thus $C$ couples the bivariate distribution, $F$, with the two marginal distributions. 
Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
$G_A$ and $G_B$ are univariate distribution functions, then 
$D(G_A(x),G_B(y))$ is a distribution function.
The applications of this study fall within the case in which univariate marginal 
distributions have continuous, strictly 
monotonic distribution functions, and this case is simpler, so we 
henceforth make such an assumption. 
Then, "the" copula (now guaranteed unique) 
associated with $(X,Y)$ is the distribution
function of the random variable $(F_X(X),F_Y(Y))$ 
(Appendix \ref{SM-ContStrictMonoton}). And 
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
vector with distribution function $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
(Appendix \ref{SM-ContStrictMonoton}). Application of the 
distributions functions of the component random variables translates
between the uniform-marginal context, which isolates the 
dependence information, and the context of the original random variable.

Thus Sklar's theorem implies that
any random vector can be constructed from a copula and marginal 
distributions. And conversely,
any copula and any univariate distribution functions give rise to a random vector. 
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables. Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure. For instance,
construction of several bivariate distributions is carried out in this way in
Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
illustrates in particular how a bivariate distribution can be changed, while retaining 
the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
is not, although its marginals
are the same. Fig. \ref{PedagogFigCopThry}D shows stronger left- than right-tail dependence. 
Sklar's theorem
also makes it possible to study and model dependence separately 
from marginal distributions. For instance, in a pedagogical 
example, @Anderson2018 modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].
<!--***DAN: dont forget to put these few words in!-->

<!--***DAN: I still feel like we have to make this more intuitive for people
at this point. Maybe a figure with two panels showing two copulas on
the left, and two panels shwoing two sets of marginals at the top,
and then four panels showing what you get when each of these combine?
Good idea. Steps: 
DONE split the current Fig 1 so the new Fig 1 has the first 6 panels,
but C-D should only have the points. 
DONE Then make a Fig 2 as above, 
DONE and edit the text in the above para to use that figure. 
STILL HAVE TO DO The rest of the panels of the current Fig 1 will make a new Fig 3 later.
-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.5in]{./Results/PedagogFigCopThry.pdf}
\caption{Normal (A) and Clayton (B) copulas were combined with standard normal 
marginals (C, D) and gamma marginals (E, F) via Sklar's theorem to produce 
bivariate distributions. Note that a normal copula is distinct from a normal marginal 
distribution: a normal copula is the copula of a bivariate normal distribution.
See the Background section for more information on normal, Clayton,
and other copulas. Each copula was used with both sets of marginals and each
set of marginals was used with both copulas, to demonstrate that: both copula and 
marginals contribute fundamentally to the resulting distribution, the copula 
contributing the information on the dependence between the variables; and that 
one can select the copula and marginals independently. Bivariate
distributions, including the copulas, are depicted via their log-scale probability density functions (pdfs), 
and marginals are depicted via their linear-scale pdfs. Grey dots are
50 random samples from each distribution. The parameter $0.7$ was used for the normal 
copula, the parameter $2$ for the Clayton 
copula, and shape and scale parameters $5$ and $1$ for the gamma marginals. 
Variables $u$ and $v$ were used for copulas
and $x$ and $y$ for distributions created by pairing a copula with
(non-uniform) marginals.
\label{PedagogFigCopThry}}
\end{center}
\end{figure}

<!--***DANr: Add a new section called Background on copulas.
It can have the following kinds of stuff (probably not exhaustive, and probably
not in order, and some of this material is currently below):
1) DONE - The definition, Sklar's theorem
2) DONE What good are copulas - intuitive idea, basically, but also have to be linked to 
the formalities of the definition
3) What some of the different copula families look like - in particular you will use BB1 in Fig S2 and cite Background, so mention it
  Three new figures: Normal, Clayton, BB1
4) What tail dependence actually is

The explanation of how ranks relate to copulas and estimating the copula parameters
How to carry out fitting (basically, MLE fitting on ranks - Genest et al paper)

You are going to have to rearrange the order of some figs in Sup Mat because now the data section comes after Bacground, in which the different copula famlies are introduced
-->

@@

We introduce some standard copula families, as example illustran

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; p is the value of the parameter for the N 
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition).\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C 
copulas. S is Spearman correlation; p is the value of the parameter for the C family (it 
is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, 
respectively (see Background section for definition).\label{figped_C}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively (see Background section for definition).\label{figped_BB1}}
\end{center}
\end{figure}

### 2019-09-10 SupportingInformation.Rmd
still working on Background section

<!--generating plots for pedagog figures-->
```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
```

@@

The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
Fig. \ref{fig_transf_pearson_spearman}A can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cummulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
the cdf of the gamma distribution with shape and scale parameters 2. 
The functions `pnorm` and `qgamma` in R were used.
Fig. \ref{fig_transf_pearson_spearman}B can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
The Pearson correlations differ substantially after transformation, 
but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
Fig. \ref{MT-fig_cop_pedag1}C (the points, not the contours) can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution.
Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption[Lower- and upper-tail dependence can vary independently of correlation]{Lower- and 
upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
copula (see section on methods for Q1 in the main text) 
with three different 
sets of parameters chosen to produce a variety of 
lower-tail (LT) and upper-tail (UT) dependence strengths, but with the same 
Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
within each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

# Proofs of statements related to Skalr's theorem in the case of continuous, strictly monotonic margins \label{ContStrictMonoton}

@@

Now let $D$ be any bivariate copula and let 
$G_A$ and $G_B$ be univariate distribution functions, so that 
$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a a distribution function.
We here show that if $G_A$ and $G_B$ and continuous and strictly monotonic, then
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
vector with uniform marginals associated with $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
The inverses exist because of the strict monotonicity assumption.
To see this, note that the distribution function of 
$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
$(U,V)$, which proves the statement. In fact, a stronger statement assuming
neither continuity nor strict monotonicity is possible, if one uses generalized
inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
statements for simplicity and because our applications are solely under the case
of continuous, strictly monotonic margins. But see @Anderson2018 and 
@Popovic2018 for applications of copulas outside this case.

@@

<!--Now the copula figures-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; p is the value of the parameter for the N 
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_C}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB1}}
\end{center}
\end{figure}

>>>>

<!--generating plots for pedagog figures-->
```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
source("./PedagogFigCopThry.R")
```

@@

The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
Fig. \ref{fig_transf_pearson_spearman}A can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cummulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
the cdf of the gamma distribution with shape and scale parameters 2. 
The functions `pnorm` and `qgamma` in R were used.
Fig. \ref{fig_transf_pearson_spearman}B can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
The Pearson correlations differ substantially after transformation, 
but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
A scatterploy very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution.
Likewise a scatterplot similar to Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption[Lower- and upper-tail dependence can vary independently of correlation]{Lower- and 
upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
copula (see Background section in the main text) 
with three different 
sets of parameters chosen to produce a variety of 
lower-tail (LT) and upper-tail (UT) dependence strengths, but with the same 
Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
within each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

# Proofs of statements related to Sklar's theorem in the case of continuous, strictly monotonic margins \label{ContStrictMonoton}

@@

Now let $D$ be any bivariate copula and let 
$G_A$ and $G_B$ be univariate distribution functions, so that 
$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a a distribution function.
We here show that if $G_A$ and $G_B$ and continuous and strictly monotonic, then
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
vector with uniform marginals associated with $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
The inverses exist because of the strict monotonicity assumption.
To get to the desired result, note that the distribution function of 
$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
$(U,V)$. In fact, a stronger statement assuming
neither continuity nor strict monotonicity is possible, if one uses generalized
inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
statements for simplicity and because our applications are solely under the case
of continuous, strictly monotonic margins. But see @Anderson2018 and 
@Popovic2018 for applications of copulas outside this case.

@@

<!--Now the copula figures that were not presented in the main text-->

@@
